Advantages of mixed quantum channels for information processing

### 2019-03-06 decorrelation/mixed_unitaries.tex
one more thing

% ==============================================================================
% Section: Generators
% ==============================================================================
\subsection{Optimization Targets}
\subsubsection{First-order generators} % (fold)
\label{sub:first_order_generators}
If our primary goal is for our MQC to have increased performance, then a useful target for constructing MQCs is to minimize the diamond norm. The diamond norm is a non-linear function that in general requires a convex optimizer to compute. However, if our errors are small enough, we can consider the linearized problem, and minimize the diamond norm to first-order. As in Section \ref{sec:simple_example}, the MQC that minimizes the diamond norm to first-order in $\epsilon$ is:
\begin{equation}\label{eq:balanced}
\mathcal{M}_{jk} = \frac{1}{2}\tr(\sigma_j^{\dagger}U_{-\epsilon}\sigma_kU_{-\epsilon}^{\dagger}) + \frac{1}{2}\tr(\sigma_j^{\dagger}U_{\epsilon}\sigma_kU_{\epsilon}^{\dagger})
\end{equation}
This channel will have a diamond norm proportional to $\epsilon^2$, whereas any other combination will have a linear dependence on $\epsilon$. Contrast this with the fact that any linear combination of $\{U_{\epsilon}, U_{-\epsilon}\}$ minimizes the AGI of the resulting channel with a value of $\epsilon^2$. While selecting the convex mixture of unitaries that minimizes the AGI is  trivial, the problem of efficiently minimizing the diamond norm can be difficult, and may be the relevant quantity to consider depending on the application. However, as discussed in \cite{Campbell2017}, a sufficient condition to minimize the diamond norm of a MQC with error generators $\{G_j\}$ to first order is:
\begin{equation}\label{eq:campbell-condition}
\sum \omega_j G_j = 0
\end{equation}
In \cite{Campbell2017} Campbell constructs an algorithm that, given an oracle to approximate unitaries, generates controls and weightings to find an MQC with this property. Alternatively one might ask the question: \textit{Given a fixed collection of controls, how can I produce weighting a weighting with minimal diamond norm?} Such a situation may arise if there is a natural family of controls that implement a desired gate, or if, as we consider later (Section \ref{sec:numerical_results}), we randomly generate a collection of gate implementations. In these situations, one can equivantly use convex optimization to solve this problem. Consider the matrix whose columns are the vectorized Hamiltonians at our disposal, i.e. for $m$, $n\times n$ Hamiltonians:
\begin{equation}\label{eq:vectorized_hamiltonians}
	\bold{G} = \left(\begin{array}{cccc}
		G_{1_{11}} & G_{2_{11}} & \ldots   \\ 
		\vdots\ & \ddots &    \\
		G_{1_{nn}} &  &  G_{m_{nn}} \\ 
	\end{array} 	
	\right)
\end{equation}
If our weighting vector for our MQC is $\omega$, we can rewrite this sum as a matrix product whose two-norm will be zero if and only if the sum is zero. Additionally, this optimization needs to be constrained so that $\omega$ only contains positive values that sum to one. This forms a convex optimization problem, and can be written as:

@@

\subsubsection{Off-diagonals} % (fold)
\label{sub:off_diagonals}
Instead of targeting the diamond norm and vectorizing the error generators, we can instead consider error channels $\error_i$ such that $\error_i\actual_i=\target$, where $\target$ is again the target gate. To approximate a depolarizing channel we may define a different minimization problem to  minimize the off-diagonal terms:
\begin{equation}\label{eq:minimization}
  \begin{split}
    &\underset{w_0, ..., w_N}{\textbf{minimize}} \{\sum_{i\neq j}^N|\sigma_i\Lambda(\sigma_j)|^2\}\\
    &\textbf{where}\ \Lambda(\sigma_j) := \sum^N_{i=1}w_i\mathcal{E}_i^{\dagger}\sigma_j\mathcal{E}_i\\
    &\textbf{subject to} \sum_{i=1}^Nw_i = 1
  \end{split}
\end{equation}

Previous authors have considered minimizing the diamond distance to the nearest Pauli or Clifford Channel \cite{Magesan2013}, and while this gives a good theoretical framework, it requires the use of a convex solver, and in particular does not have the restriction that the optimal channel be decomposable into a given family of controls. Therefore we can view our approach as a regularization of this method, where we have imposed the condition that our final channel must arise from a BCS. In doing so, we move from computing the diamond norm to needing to computing a sum. In the next section we give a simple example, followed by numerical results for one qubit and two qubits gates.
% subsection off_diagonals (end)

% ==============================================================================
% Section: Robustness
% ==============================================================================
\subsection{Additional Constraints}
\subsubsection{Adding robustness} % (fold)
\label{sub:adding_robustness}
\note{ Try to make this more readable }
While mixed quantum channels offer significant improvements to gate performance, they fail to take into account the reality that most control electronics experience drift over time scales relevant to QIP performance. Because of this drift, the quality of the MQC will degrade. Thus, we would like to design MQCs that are \textit{robust} to this drift. To enforce robustness, we can consider higher derivatives of the generators in Equation \ref{eq:campbell-condition}. Instead of only requiring that the $0^{th}$-order derivatives average to zero, we will impose a similar condition on the higher-order derivatives with respect to parameters that may drift:
\begin{equation}
D^n_j = \frac{1}{n!}\frac{\partial^{n}}{\partial\delta_{i_1}\ldots\partial\delta_{i_n}}H_j(\vec{\delta})|_{\delta=\vec{0}}
\end{equation}
If the dimension of $\vec{\delta}$ is $d$, the indices $i_0, \ldots, i_n$ take on values in $1, \ldots, d$ and this matrix has $d^n$ entries. 
We say that a mixed quantum channel is robust to order $\ell$ (an $\ell$MQC) if for all $1 \leq k \leq \ell$:
\begin{equation}\label{eq:MQC}
\begin{gathered}
\sum_j\omega_j(\sum_{n=0}^k D^n_j)^n = \vec{0}\\
\end{gathered}
\end{equation}
In particular, we see that a 0MQC satisfies Equation \ref{eq:campbell-condition}. More generally, these conditions imply that an $\ell$MQC is insensitive to the $\ell^{th}$ order in drift in $\vec{\delta}$. To see this, we can rewrite the error on each control in the MQC as:
\begin{equation}\label{eq:taylor}
\begin{gathered}
\actual_j(\vec{\delta}) = \exp(-i(H_j(\vec{0}) + \frac{\partial}{\partial\delta_i}H_j(d\delta_i)\\ +  \frac{1}{2}\frac{\partial^2}{\partial\delta_i\partial\delta_k} H_j(d\delta_i d\delta_k) + \ldots))\target
\end{gathered}
\end{equation}
By Taylor expanding Equation \ref{eq:taylor} in $\vec{\delta}$, one finds that the the first $\ell$ derivatives of an $\ell$MQC will be zero. Furthermore, if we are only interested in being first order insensitve to drift and can find controls such that $|D_j^n|\approx\epsilon$, we can approximate Equation \ref{eq:MQC} as:
\begin{equation}\label{eq:MQC-relaxed}
\begin{gathered}
\sum\omega_jD^n_j = 0\\
\end{gathered}
\end{equation}
This condition guarantees that errors will be supressed quadratically for all derivatives up to order $\ell$. A proof is included in the appendix that generalizes the Hastings-Campbell Mixing Lemma in \cite{Campbell2017}. Namely, Campbell showed that if $0\in $ Conv$[\{H_i(\vec{\delta})\}]$, where Conv is the convex hull of its arguments, then an $\vec{\omega}$ exists that quadratically decreases the diamond norm. We prove that $0\in $ Conv$[\{D_j^n(\vec{\delta})\}]$ implies there is an $\vec{\omega}$ exists that quadratically decreases the $\ell^{th}$-order sensitivity of the diamond norm of an $\ell$MQC. Figure \ref{fig:vectorspace} gives geometric intuition for the conditions required to produce an $\ell$MQC.

To generate robustly mixed quantum channels, we first define the vectorized derivative matrix ${D^{\ell}}$ in a similar way to Equation \ref{eq:vectorized_hamiltonians}:
\begin{equation}
{\bold{D}^{\ell}} =  \left(\begin{array}{cccc}
		D^\ell_{1_{11}} & D^\ell_{2_{21}} & \ldots   \\ 
		\vdots\ & \ddots &    \\
		D^\ell_{m_{11}} &  &  D^\ell_{m_{nn}} \\ 
	\end{array} 	
	\right)
\end{equation}

@@

\begin{equation}\label{eq:robust_minimization}
  \begin{split}
    &\underset{\omega_j\geq0, |\omega|_1=1}{\textbf{minimize}: } ||{\bold{D}^{\ell}}^T\omega||\\
    &\textbf{subject to: } \forall n<\ell, \sum \omega_jD_j^n=\vec{0}\\
  \end{split}
\end{equation}
with $D_n^j$ defined in Equation \ref{eq:taylor}. The time to solve these convex optimization problems is independent of the number of rows, which in our case is the number of drifting parameters. However, the size of the matrix that must be computed before solving grows as $Nd^{\ell}$, and the runtime of solving the optimization problem grows as $N^4$, with $d$ being the number of drifting parameters, and $N$ being the number of controls.

@@

\begin{equation}\label{eq:minimization_l2}
\begin{split}
&\underset{n\in[N]}{\textbf{minimize}}\{\\
&\ \ \bunderbrace{\textbf{minimize}: }{\omega_j\geq0, |\omega|_1=1} ||{\bold{D}^{\ell}}^T\omega||+ \eta\sum\omega_j||D^0_j||_2\\
&\ \ \textbf{subject to: } \forall n<\ell, \sum \omega_jD_j^n=0\\
&\}
\end{split}
\end{equation}
with $\eta \geq 0$. By making $\eta$ larger, we can supress the diamond norm to second order, while ensuring that we are selecting those that will give us a smaller prefactor. The absolute value of $\eta$ depends on the particular numerical values in $\bold{D}^\ell$. Naively generating the 1Q 0MQC in the previous section results in nontrivial support on all the members of the control family. However, by rewriting the minimization to impose this sparsity constraint discussed in Section \ref{sec:robustly_mixed}, the resulting 0MQC uses just five of the controls. This shows that through adding constraints to our optimization routine, we can make the MQC practically useful.

@@

\begin{equation}\label{eq:minimization_regularization}
\begin{split}
&\underset{n\in[N]}{\textbf{minimize}}\{\\
&\ \ \bunderbrace{\textbf{minimize}: }{\omega_j\geq0, |\omega|_1=1,\\ t\geq0} ||{\bold{D}^{\ell}}^T\omega|| + t\\
&\ \ \textbf{subject to: } \omega_n > \frac{\lambda}{t}\\
&\ \ \phantom{\textbf{subject to: }} \forall n<\ell, \sum \omega_jD_j^n=0\\
&\}
\end{split}
\end{equation} with $\lambda\geq0$. As with $\eta$ in the last subsection, the optimal value of this parameter is problem specific, depending on $\bold{D}^\ell$, and how sparse the solution needs to be. For $\lambda$ small, the problem reduces to the original problem, and for $\lambda$ large we see that if $\omega$ is not sparse then $t$ must be very large, which increases the cost of that particular solution.
In both the 0MQC and the 1MQC case in our numerical implementation we imposed the same $\ell_2$ penalty, so that the algorithm preferentially selects controls with smaller errors. Adding this constrain made the 0MQC perform better at the origin by nearly an order of magnitude, and moved the 1MQC from being out-performed by the 0MQC for all detunings, to out performing by nearly an order of magnitude when there is $.1\%$ drift in the controls. Imposing this constraint allows us to trade off flatness at the origin for performance.

>>>>

% ==============================================================================
% Section: Generators
% ==============================================================================
\subsection{Optimization Targets}
\subsubsection{First-order generators} % (fold)
\label{sub:first_order_generators}
If our primary goal is for our MQC to have increased performance, then a useful target for constructing MQCs is to minimize the diamond norm. The diamond norm is a non-linear function that in general requires a convex optimizer to compute. However, if our errors are small enough, we can consider the linearized problem, and minimize the diamond norm to first-order. As in Section \ref{sec:simple_example}, the MQC that minimizes the diamond norm to first-order in $\epsilon$ is:
\begin{equation}\label{eq:balanced}
\mathcal{M}_{jk} = \frac{1}{2}\tr(\sigma_j^{\dagger}U_{-\epsilon}\sigma_kU_{-\epsilon}^{\dagger}) + \frac{1}{2}\tr(\sigma_j^{\dagger}U_{\epsilon}\sigma_kU_{\epsilon}^{\dagger})
\end{equation}
This channel will have a diamond norm proportional to $\epsilon^2$, whereas any other combination will have a linear dependence on $\epsilon$. Contrast this with the fact that any linear combination of $\{U_{\epsilon}, U_{-\epsilon}\}$ minimizes the AGI of the resulting channel with a value of $\epsilon^2$. While selecting the convex mixture of unitaries that minimizes the AGI is  trivial, the problem of efficiently minimizing the diamond norm can be difficult, and may be the relevant quantity to consider depending on the application. However, as discussed in \cite{Campbell2017}, a sufficient condition to minimize the diamond norm of a MQC with error generators $\{G_j\}$ to first order is:
\begin{equation}\label{eq:campbell-condition}
\sum \omega_j G_j = 0
\end{equation}
In \cite{Campbell2017} Campbell constructs an algorithm that, given an oracle to approximate unitaries, generates controls and weightings to find an MQC with this property. Alternatively one might ask the question: \textit{Given a fixed collection of controls, how can I produce weighting a weighting with minimal diamond norm?} Such a situation may arise if there is a natural family of controls that implement a desired gate, or if, as we consider later (Section \ref{sec:numerical_results}), we randomly generate a collection of gate implementations. In these situations, one can equivantly use convex optimization to solve this problem. Consider the matrix whose columns are the vectorized Hamiltonians at our disposal, i.e. for $m$, $n\times n$ Hamiltonians:
\begin{equation}\label{eq:vectorized_hamiltonians}
	\bold{G} = \left(\begin{array}{cccc}
		G_{1_{1,1}} & G_{2_{1,1}} & \ldots   \\ 
		\vdots\ & \ddots &    \\
		G_{1_{n,n}} &  &  G_{m_{n,n}} \\ 
	\end{array} 	
	\right)
\end{equation}
If our weighting vector for our MQC is $\omega$, we can rewrite this sum as a matrix product whose two-norm will be zero if and only if the sum is zero. Additionally, this optimization needs to be constrained so that $\omega$ only contains positive values that sum to one. This forms a convex optimization problem, and can be written as:

@@

\subsubsection{Off-diagonals} % (fold)
\label{sub:off_diagonals}
Instead of targeting the diamond norm and vectorizing the error generators, we can instead consider error channels $\error_i$ such that $\error_i\actual_i=\target$, where $\target$ is again the target gate. To approximate a depolarizing channel we may define a different minimization problem to  minimize the off-diagonal terms. Consider the matrix of off-diagonal terms in Equation \ref{eq:process_matrix}:

\begin{equation}\label{eq:vectorized_off_diagonal}
	\bold{R} = \left(\begin{array}{cccc}
		R_{1_{1,1}} & R_{2_{1,1}} & \ldots   \\ 
		\vdots\ & \ddots &    \\
		R_{1_{n-1,n-1}} &  &  R_{m_{n-1,n-1}} \\ 
	\end{array} 	
	\right)
\end{equation}
then again we consider the same minimziation problem, where we replace $\bold{G}$ with $\bold{R}$:

\begin{equation}\label{eq:minimization_diagonal}
  \begin{split}
    &\underset{\omega_j\geq0, |\omega|_1=1}{\textbf{minimize}: } ||\bold{R}\omega||_2\\
  \end{split}
\end{equation}

While previous authors have considered minimizing the diamond distance to the nearest Pauli or Clifford Channel \cite{Magesan2013}, this requires many evaluations of the diamond norm, and in particular does not have the restriction that the optimal channel be decomposable into a given family of controls. Our approach, on the other hand, guarantees that the resulting MQC is able to be implemented with the available controls, and is an easy-to-solve quadratic program.

Depending on which properties are desired of a MQC, many different optimization targets can be chosen. From the examples given above, it is clear that if one desires to produce a channel that maximizes worst-case performance, considering the first-order generators is an appropriate target. If instead the goal is produce a MQC whose error channel is more tractable to simulate, minimzing the off-diagonal elements may be more appropriate.
% subsection off_diagonals (end)

% ==============================================================================
% Section: Robustness
% ==============================================================================
\subsection{Additional Constraints}
\subsubsection{Adding robustness} % (fold)
\label{sub:adding_robustness}
While mixed quantum channels offer significant improvements to gate performance, they fail to take into account the reality that most control electronics experience drift over time scales relevant to QIP performance. Because of this drift, the quality of the MQC will degrade. Thus, we would like to design MQCs that are robust to this drift. To enforce robustness, we can consider higher derivatives of the generators in Equation \ref{eq:campbell-condition}. Instead of only requiring that the error generators average to zero, we will impose a similar condition on the higher-order derivatives with respect to parameters that may drift. If our control Hamiltonian is parameterized by a vector $\vec{\delta}$, then the quantity we will be interested in is:
\begin{equation}
D^n_j = \frac{1}{n!}\frac{\partial^{n}}{\partial\delta_{i_1}\ldots\partial\delta_{i_n}}H_j(\vec{\delta})|_{\delta=\vec{0}}
\end{equation}
If the dimension of $\vec{\delta}$ is $d$, the indices $i_0, \ldots, i_n$ take on values in $1, \ldots, d$ and this matrix has $d^n$ entries. 
We say that a mixed quantum channel is robust to order $\ell$ (an $\ell$MQC) if for all $1 \leq k \leq \ell$:
\begin{equation}\label{eq:MQC}
\begin{gathered}
\sum_j\omega_j(\sum_{n=0}^k D^n_j)^n = \vec{0}\\
\end{gathered}
\end{equation}
In particular, we see that a 0MQC satisfies Equation \ref{eq:campbell-condition}. More generally, these conditions imply that an $\ell$MQC is insensitive to the $\ell^{th}$ order in drift in $\vec{\delta}$. To see this, we can rewrite the error on each control in the MQC as:
\begin{equation}\label{eq:taylor}
\begin{gathered}
\actual_j(\vec{\delta}) = \exp(-i(H_j(\vec{0}) + \frac{\partial}{\partial\delta_i}H_j(d\delta_i)\\ +  \frac{1}{2}\frac{\partial^2}{\partial\delta_i\partial\delta_k} H_j(d\delta_i d\delta_k) + \ldots))\target
\end{gathered}
\end{equation}
By Taylor expanding Equation \ref{eq:taylor} in $\vec{\delta}$, one finds that the the first $\ell$ derivatives of an $\ell$MQC will be zero. Furthermore, if we are only interested in being first order insensitve to drift and can find controls such that $|D_j^n|\approx\epsilon$, we can approximate Equation \ref{eq:MQC} as:
\begin{equation}\label{eq:MQC-relaxed}
\begin{gathered}
\sum\omega_jD^n_j = 0\\
\end{gathered}
\end{equation}
This condition guarantees that errors will be supressed quadratically for all derivatives up to order $\ell$. A proof is included in the appendix that generalizes the Hastings-Campbell Mixing Lemma in \cite{Campbell2017}. Namely, Campbell showed that if $0\in $ Conv$[\{H_i(\vec{\delta})\}]$, where Conv is the convex hull of its arguments, then an $\vec{\omega}$ exists that quadratically decreases the diamond norm. We prove that $0\in $ Conv$[\{D_j^n(\vec{\delta})\}]$ implies there is an $\vec{\omega}$ exists that quadratically decreases the $\ell^{th}$-order sensitivity of the diamond norm of an $\ell$MQC. Figure \ref{fig:vectorspace} gives geometric intuition for the conditions required to produce an $\ell$MQC.

To generate robustly mixed quantum channels, we first define the vectorized derivative matrix ${D^{\ell}}$ in a similar way to Equation \ref{eq:vectorized_hamiltonians}:
\begin{equation}
{\bold{D}^{\ell}} =  \left(\begin{array}{cccc}
		D^\ell_{1_{1,1}} & D^\ell_{2_{1,1}} & \ldots   \\ 
		\vdots\ & \ddots &    \\
		D^\ell_{1_{n,n}} &  &  D^\ell_{m_{n,n}} \\ 
	\end{array} 	
	\right)
\end{equation}

@@

\begin{equation}\label{eq:robust_minimization}
  \begin{split}
    &\underset{\omega_j\geq0, |\omega|_1=1}{\textbf{minimize}: } ||{\bold{D}^{\ell}}\omega||\\
    &\textbf{subject to: } \forall n<\ell, \sum \omega_jD_j^n=\vec{0}\\
  \end{split}
\end{equation}
with $D_n^j$ defined in Equation \ref{eq:taylor}. The time to solve these convex optimization problems is independent of the number of rows, which in our case is the number of drifting parameters. However, the size of the matrix that must be computed before solving grows as $Nd^{\ell}$, and the runtime of solving the optimization problem grows as $N^4$, with $d$ being the number of drifting parameters, and $N$ being the number of controls.

@@

\begin{equation}\label{eq:minimization_l2}
\begin{split}
&\underset{n\in[N]}{\textbf{minimize}}\{\\
&\ \ \bunderbrace{\textbf{minimize}: }{\omega_j\geq0, |\omega|_1=1} ||{\bold{D}^{\ell}}\omega||+ \eta\sum\omega_j||D^0_j||_2\\
&\ \ \textbf{subject to: } \forall n<\ell, \sum \omega_jD_j^n=0\\
&\}
\end{split}
\end{equation}
with $\eta \geq 0$. By making $\eta$ larger, we can supress the diamond norm to second order, while ensuring that we are selecting those that will give us a smaller prefactor. The absolute value of $\eta$ depends on the particular numerical values in $\bold{D}^\ell$. Naively generating the 1Q 0MQC in the previous section results in nontrivial support on all the members of the control family. However, by rewriting the minimization to impose this sparsity constraint discussed in Section \ref{sec:robustly_mixed}, the resulting 0MQC uses just five of the controls. This shows that through adding constraints to our optimization routine, we can make the MQC practically useful.

@@

\begin{equation}\label{eq:minimization_regularization}
\begin{split}
&\underset{n\in[N]}{\textbf{minimize}}\{\\
&\ \ \bunderbrace{\textbf{minimize}: }{\omega_j\geq0, |\omega|_1=1,\\ t\geq0} ||{\bold{D}^{\ell}}\omega|| + t\\
&\ \ \textbf{subject to: } \omega_n > \frac{\lambda}{t}\\
&\ \ \phantom{\textbf{subject to: }} \forall n<\ell, \sum \omega_jD_j^n=0\\
&\}
\end{split}
\end{equation} with $\lambda\geq0$. As with $\eta$ in the last subsection, the optimal value of this parameter is problem specific, depending on $\bold{D}^\ell$, and how sparse the solution needs to be. For $\lambda$ small, the problem reduces to the original problem, and for $\lambda$ large we see that if $\omega$ is not sparse then $t$ must be very large, which increases the cost of that particular solution.
In both the 0MQC and the 1MQC case in our numerical implementation we imposed the same $\ell_2$ penalty, so that the algorithm preferentially selects controls with smaller errors. Adding this constrain made the 0MQC perform better at the origin by nearly an order of magnitude, and moved the 1MQC from being out-performed by the 0MQC for all detunings, to out performing by nearly an order of magnitude when there is $.1\%$ drift in the controls. Imposing this constraint allows us to trade off flatness at the origin for performance.
